Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-13 ms.tex
Polished general intro and shortened previous works, also added recent neural network papers to .bib.  Still need to incorporate these in intro

\begin{document} 
\shorttitle{Automated Lensing Learner}
\shortauthors{Avestruz, Li, Lightman}
\submitted{The Astrophysical Journal} 
\slugcomment{The Astrophysical Journal, submitted}

@@

\author{Camille Avestruz\altaffilmark{1-3,$\dagger$}\thanks{E-mail:
    avestruz@uchicago.edu}, Nan Li\altaffilmark{2-4}, and Matthew Lightman\altaffilmark{5} }

\affil{
$^1${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^2${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^3${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.};\\
$^4${High Energy Physics Division, Argonne National Laboratory,
Lemont, IL 60439};\\
$^5${Digital Intelligence, JPMorgan Chase, Chicago, IL 60603 U.S.A.};\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
  \href{mailto:avestruz@uchicago.edu}{avestruz@uchicago.edu}\\
}

@@

% Gravitational lensing as a geometric (?) test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  Strong gravitational
lensing can manifest as visible giant arcs magnifying high redshift
galaxies \citep{lyndsandpetrosian_86,gladders_etal03}, multiply imaged
quasars \citep{walsh_etal79}, and arclets.  Lensing signatures probe
the underlying dark matter distribution of the lens
\citep{warrenanddye_03}, high redshift galaxy formation
\citep{allam_etal07}, and provide a geometric test of cosmology via
comparison of predicted arc abundances and observed abundances
\citep{kochanek_96,chae_03,linder_04} and time-delay signals from
multiply imaged quasars ???\citep{}.

@@

% Arc identification can be done visually (e.g. that zooniverse
% thing).  But, in the advent of optical surveys, DES and LSST, we
% will be limited by our ability to automatically process the large
% amounts of data.  Wide field surveys - SPT in millimeter, Herschel
% in submillimeter; high resolution spectroscopic imaging with ALMA;
% optical surveys DES, HSC, LSST, and Euclid (Oguri \& Marshall 2010).
Visual arc identification has been effective through the use of
citizen science platforms (e.g. \citet{marshall_etal16,more_etal16}).
However, in the advent of large surveys, such as the Dark Energy
Survey (DES) and the Large Synoptic Sky Telescope (LSST), we will be
limited by our ability to automatically process large amounts of data
and select likely strong lensing candidates for follow up visual and
spectroscopic confirmation.  The strong lensing sector requires an
exploration of techniques in image processing and classification.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  {\em
  SpaceWarps} is an example of citizen science based image
classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  Visual classification has been
quite successful for a dataset like CFHTLS, classifying \change{XX in
  XX time}.  However, future datasets like Euclid and LSST will be
\change{XX} times this size, challenging the scalability of a citizen
science approach.  The next step is to find automated methods with
performance comparable to or better than humans. {\em SpaceWarps} is a
part of the Zooniverse Project (CITE???), which also includes {\em
  Galaxy Zoo}, the citizen science based image classification of
galaxy types \citep{lintott_etal08}.  Galaxy classification is a
successful example of in which machine learning algorithms
successfully train models to classify images with comparable
performance to humans \citep{dieleman_etal15}.

% Describe general robot identification schemes that have been
% implemented
General automated, or ``robot'', identification refers to a classifier
where images are fed into an image processing pipeline that enhances
and extracts characteristic features as parameters on which a pattern
recognition piece is applied.  Examples of this include ArcFinder for
group or cluster scale lens \citep{seidelandbartelmann_07}, Ringfinder
for multiply imaged quasars\citep{gavazzi_etal14}, and ???.  Each
pipeline .... ... and relies on the image morphology where ... ...
Note, that the method best applied to ... ... and the objects of
interest.

@@

% Previous works and our goal to make a universal pipeline 
Machine learning to automate.  Previous works (StrongML),
More+... Here, we focus on the improvement from preprocessing the
images and describe the pipeline, which is publicly available.
Depending on the image type, different preprocessing is likely
necessary (rings, multiply lensed quasars, etc.) Our goal is to create
an open source pipeline that is general enough for the user to select
or add appropriate image processing techniques to augment the features
of the lensed image, train and test data with different machine
learning techniques, and to select the sequence of methods that
maximizes completeness with optimal purity.

@@

%  Relevance of our work
While there has been a recent surge in the use of neural networks
applied to image classification problems, these implementations are
not easily scalable, nor easily accessible to the entire scientific
community.  We present a supervised classification pipeline built with
open source tools, tested on mock HST and LSST data.  Results are
reproducible on personal computers, as both the pipeline and the data
will be publicly distributed.

@@

We use mock Hubble Space Telescope (HST) and Large Synoptic Sky
Telescope (LSST) images generated with PICS (Pipeline for Images of
Cosmological Strong lensing) (Li et al. 2016) and LensPOP (Thomas
Colletti 2013) as training set and testing set. In these images, the
lensed images include the images of lens galaxies, lensed images, and
galaxies on the line of sight; the unlensed images include the images
of lens galaxies and the galaxies on the line of sight.

@@

% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope} (HST) and {\em Large Synoptic
  Sky Telescope} (LSST) images generated with PICS (Pipeline for
Images of Cosmological Strong lensing) \citep{li_etal16} and
LensPop\citep{collett_etal15}.  These images include cluster-member
galaxies and foreground stars.  Half of the images used are strong
lensing systems.  Here, lensed galaxy images are ray-traced images of
actual galaxies extracted from deep Hubble Space Telescope
observations.  The final mock images are ``observed'' with a realistic
point spread function corresponding to modeled detector artifacts for
bright stars.  See \cite{li_etal16} for more details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fp.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bn.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fn.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tn.pdf}\\
\caption{LSST 10 year mock images. Left two columns: Lens containing
  images, annotated with the image score assigned by our trained
  classifier.  Right two columns: Non-lens containing images,
  annotated with the image score.  The top two rows show
  characteristic images that will be accepted with a high threshold
  for classification, contributing to the bottom left of the ROC curve
  in Figure~\ref{fig:ROCcompilation}.  The middle two rows show
  characteristic images that will be accepted with a moderate
  threshold, contributing to the knee of the ROC curve, with true
  positive rates and false positive rates of $tpr\approx0.8$ and
  $fpr\approx0.2$, respectively.  The bottom two rows show
  characteristic images that will only be accepted with an extremely
  lenient threshold, contributing to the top right area of the ROC
  curve.}\label{fig:ROCsamplesLSST10}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Borderlines
The middle two rows of Figure~\ref{fig:ROCsamplesLSST10} illustrate
the borderline positives and borderline negatives. The successful
classification of the borderline positives and negatives are most
sensitive to the threshold, and would be the first candidates for
alternative classification methods, such as visual follow-up.
Thresholds set around these scores yield a true positive rate and
false positive rate of $tpr\approx0.8$ and $fpr\approx0.2$,
respectively.

>>>>

\begin{document} 
\shorttitle{Automated Lensing Learner}
\shortauthors{Avestruz, et al.}
\submitted{The Astrophysical Journal} 
\slugcomment{The Astrophysical Journal, submitted}

@@

\author{Camille Avestruz\altaffilmark{1-3,$\dagger$}\thanks{E-mail:
    avestruz@uchicago.edu}, Nan Li\altaffilmark{2-4}, Matthew Lightman\altaffilmark{5}, Thomas E.~Collett\altaffilmark{6}}

\affil{
$^1${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^2${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^3${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.};\\
$^4${High Energy Physics Division, Argonne National Laboratory,
Lemont, IL 60439};\\
$^5${Digital Intelligence, JPMorgan Chase, Chicago, IL 60603 U.S.A.};\\
$^6${Institute of Cosmology and Gravitation, University of Portsmouth, Burnaby Rd, Portsmouth, PO1 3FX, UK};\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
  \href{mailto:avestruz@uchicago.edu}{avestruz@uchicago.edu}\\
}

@@

% Gravitational lensing as a geometric test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  Strong gravitational
lensing can manifest as visible giant arcs magnifying high redshift
galaxies \citep{lyndsandpetrosian_86,gladders_etal03}, multiply imaged
quasars \citep{walsh_etal79}, multiply imaged galaxies
\citep{sharon_etal05}, and arclets \citep{bezecourt_etal98}.  Lensing
signatures probe the underlying dark matter distribution of the lens
\citep{warrenanddye_03}, high redshift galaxy formation
\citep{allam_etal07}, and provide a geometric test of cosmology via
comparison of predicted arc abundances and observed abundances
\citep{kochanek_96,chae_03,linder_04} and time-delay signals from
multiply imaged quasars\citep{xli_etal12,suyu_etal14,suyu_etal16}.

@@

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  Visual arc
identification has been effective through the use of citizen science
platforms.  {\em SpaceWarps} is an example of citizen science based
image classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  These platforms are quite
successful for a dataset like CFHTLS, classifying \todo{XX in XX
  time}\citep{}.  However, future datasets like Euclid
\citep{oguriandmarshall_10} and LSST \cite{} will be \todo{XX} times
this size, challenging the scalability of a citizen science approach.

% Transition to automated identification
Recent efforts have focused on the development of automated methods
with performance comparable to or better than humans. {\em SpaceWarps}
is a part of the Zooniverse Project \citep{marshall_etal16}, which
also includes {\em Galaxy Zoo}, the citizen science based image
classification of galaxy types \citep{lintott_etal08}.  Galaxy
classification is an early successful example of in which machine
learning algorithms successfully train models to classify astronomical
images with comparable performance to humans \citep{dieleman_etal15}.

% Early efforts
Earlier efforts on automated, or ``robot'', identification of strong
lensing systems have two distinct steps.  The first enhances and
extracts characteristic features, and the second uses some form of
pattern recognition to classify lens and non-lens containing systems.
{\em ArcFinder} is one such code that finds arcs in groups or cluster
scale lens \citep{seidelandbartelmann_07}, and {\em RingFinder} an
analogous tool in searching for multiply imaged quasars
\citep{gavazzi_etal14}.

% Recent works, deep learning
Much more recent works have made use of neural networks to

@@

% Previous works and our goal to make a universal pipeline 
Machine learning to automate.  Previous works \citep{lenzen_etal04},
(StrongML), More+... Here, we focus on the improvement from
preprocessing the images and describe the pipeline, which is publicly
available.  Depending on the image type, different preprocessing is
likely necessary (rings, multiply lensed quasars, etc.) Our goal is to
create an open source pipeline that is general enough for the user to
select or add appropriate image processing techniques to augment the
features of the lensed image, train and test data with different
machine learning techniques, and to select the sequence of methods
that maximizes completeness with optimal purity.

@@

%  Relevance of our work
While there has been a recent surge in the use of neural networks
applied to image classification problems, \wording{these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.}  We present a supervised
classification pipeline built with open source tools, tested on mock
HST and LSST data.  Results are reproducible on personal computers, as
both the pipeline and the data will be publicly distributed.  This
first paper serves as a case study for one feature extraction and
classical machine learning method, respectively, the {\em histogram of
  oriented gradients} and {\em logistic regression}.  This framework
is easily extendable to other feature methods and machine learning
algorithms for supervised classification.

@@

We use mock Hubble Space Telescope (HST) and Large Synoptic Sky
Telescope (LSST) images generated with PICS (Pipeline for Images of
Cosmological Strong lensing) \citep{li_etal16} and LensPOP
\citep{collett_15} as training set and testing set. In these images,
the lensed images include the images of lens galaxies, lensed images,
and galaxies on the line of sight; the unlensed images include the
images of lens galaxies and the galaxies on the line of sight.

@@

% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope} (HST) and {\em Large Synoptic
  Sky Telescope} (LSST) images generated with PICS (Pipeline for
Images of Cosmological Strong lensing) \citep{li_etal16} and
LensPop\citep{collett_15}.  These images include cluster-member
galaxies and foreground stars.  Half of the images used are strong
lensing systems.  Here, lensed galaxy images are ray-traced images of
actual galaxies extracted from deep Hubble Space Telescope
observations.  The final mock images are ``observed'' with a realistic
point spread function corresponding to modeled detector artifacts for
bright stars.  See \cite{li_etal16} for more details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fp.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bn.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fn.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tn.pdf}\\
\caption{LSST 10 year mock images. Left two columns: Lens containing
  images, annotated with the image score assigned by our trained
  classifier.  Right two columns: Non-lens containing images,
  annotated with the image score.  The top two rows show
  characteristic images that will be accepted with a high threshold
  for classification, contributing to the bottom left of the ROC curve
  in Figure~\ref{fig:ROCcompilation}.  The middle two rows show
  characteristic images that will be accepted with a moderate
  threshold, contributing to the knee of the ROC curve, with true
  positive rates and false positive rates of $tpr\approx0.8$ and
  $fpr\approx0.25$, respectively.  The bottom two rows show
  characteristic images that will only be accepted with an extremely
  lenient threshold, contributing to the top right area of the ROC
  curve.}\label{fig:ROCsamplesLSST10}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Borderlines
The middle two rows of Figure~\ref{fig:ROCsamplesLSST10} illustrate
the borderline positives and borderline negatives. The successful
classification of the borderline positives and negatives are most
sensitive to the threshold, and would be the first candidates for
alternative classification methods, such as visual follow-up.
Thresholds set around these scores yield a true positive rate and
false positive rate of $tpr\approx0.8$ and $fpr\approx0.25$,
respectively.
